The Castro AMR Simulation Code: Current and Future Developments

### 2019-10-01 astronum_2019_Castro/astronum-castro-2019.tex
more work

In modeling stellar explosions, it is often the case that the hydrodynamics method
and the nuclear reaction network require different timesteps in order to produce stable
and accurate models.  For the stiff nuclear reactions we encounter in stars, implicit ODE
method are often used together with operator splitting to evolve the nuclear reactions separately
from the hydrodynamics.  In the limit of small timesteps, the two processes are well-coupled together, but 
The recent focus in \castro\ has been on high-fidelity simulations of
reactive flow.  In \cite{castro:sdc} we introduced a new time
integration strategy, spectral deferred corrections (SDC), that
eliminates the coupling error introduced by commonly used operator
splitting techniques (see the discussion in \cite{astronum:2018} for a
graphical illustration of splitting error).  The SDC algorithm used in
\castro\ follows the ideas of \cite{dutt:2000,minion:2003}, and uses
low order explicit advection and implicit reaction updates in a
correction equation that when applied iteratively achieves high-order
time-accuracy.

In an operator split implementation, reaction and burning proceed
without knowing about the other process.  In contrast, the update in
the SDC formulation explicitly couples the processes together.  In \castro\ we consider
reactions and hydrodynamics, and write our conservative system as:
\begin{equation}
\Uc_t = \Advs{\Uc} + \Rbs{\Uc}
\end{equation}
where $\Uc$ is the vector of conserved quantities, $\Advs{\Uc}$ is the
advective term (the divergence of the hydrodynamic fluxes along with
hydrodynamic sources), and $\Rbs{\Uc}$ are the reactive source terms.
The update is done at specific temporal nodes, the number of location
of which are picked to give the desired temporal accuracy.  The update
from one time node $m$ to $m+1$ appears as:
    \begin{align}
      \label{eq:sdc:general}
      \avg{\Uc}^{{m+1},(k+1)} = \avg{\Uc}^{m,(k+1)}
            &+ \delta t_m \left [ \avg{\Advs{\Uc}}^{m,(k+1)} - \avg{\Advs{\Uc}}^{m,(k)} \right ] \nonumber \\
            &+ \delta t_m \left [ \avg{\Rbs{\Uc}}^{{m+1},(k+1)} - \avg{\Rbs{\Uc}}^{{m+1},(k)} \right ] \nonumber \\
            &+ \int_{t^m}^{t^{m+1}} dt  \left (\avg{\Advs{\Uc}}^{(k)} + \avg{\Rbs{\Uc}}^{(k)}\right)
   \end{align}
We use a finite-volume formalism, so $\avg{\Uc}$ is the average of the
conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the average
advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{k+1}$ is
the average of the reactive source at time $t^{m+1}$.  The second
superscript on each term, $(k)$ or $(k+1)$, is the iteration.  This
update is an implicit equation for the new state,
$\avg{\Uc}^{{m+1},(k+1)}$.  The last term in the update is an integral
over the sources constructed from the previous iteration's values at
each time node.  Each iteration of the SDC method increases the
temporal order of accuracy by one, up to the order of accuracy with
which the integral is constructed.  \castro\ implements both a
second-order method (using a trapezoid rule for the integral) and
fourth-order method in space and time (using a Simpsons rule for the
integral and the spatial reconstruction of
\cite{mccorquodalecolella}).  At the moment, the method is limited to
single levels.

>>>>

In modeling stellar explosions, it is often the case that the
hydrodynamics method and the nuclear reaction network require
different timesteps in order to produce stable and accurate models.
For the stiff nuclear reactions we encounter in stars, implicit ODE
method are often used together with operator splitting to evolve the
nuclear reactions separately from the hydrodynamics.  In the limit of
small timesteps, the two processes are well-coupled together, but
these conditions are not always met in simulations.  The recent focus
in \castro\ has been on high-fidelity simulations of reactive flow.
In \cite{castro:sdc} we introduced a new time integration strategy,
spectral deferred corrections (SDC), that eliminates the coupling
error introduced by commonly used operator splitting techniques (see
the discussion in \cite{astronum:2018} for a graphical illustration of
splitting error).

The SDC algorithm used in \castro\ follows the ideas of
\cite{dutt:2000,minion:2003}, and uses low order explicit advection
and implicit reaction updates in a correction equation that when
applied iteratively achieves high-order time-accuracy.  In an operator
split implementation, reaction and burning proceed without knowing
about the other process.  In contrast, the update in the SDC
formulation explicitly couples the processes together.  In \castro\ we
consider reactions and hydrodynamics, and write our conservative
system as:
\begin{equation}
\Uc_t = \Advs{\Uc} + \Rbs{\Uc}
\end{equation}
where $\Uc$ is the vector of conserved quantities, $\Advs{\Uc}$ is the
advective term (the divergence of the hydrodynamic fluxes along with
hydrodynamic sources), and $\Rbs{\Uc}$ are the reactive source terms.
The update is done at specific temporal nodes, the number of location
of which are picked to give the desired temporal accuracy.  The update
from one time node $m$ to $m+1$ appears as:
    \begin{align}
      \label{eq:sdc:general}
      \avg{\Uc}^{{m+1},(k+1)} = \avg{\Uc}^{m,(k+1)}
            &+ \delta t_m \left [ \avg{\Advs{\Uc}}^{m,(k+1)} - \avg{\Advs{\Uc}}^{m,(k)} \right ] \nonumber \\
            &+ \delta t_m \left [ \avg{\Rbs{\Uc}}^{{m+1},(k+1)} - \avg{\Rbs{\Uc}}^{{m+1},(k)} \right ] \nonumber \\
            &+ \int_{t^m}^{t^{m+1}} dt  \left (\avg{\Advs{\Uc}}^{(k)} + \avg{\Rbs{\Uc}}^{(k)}\right)
   \end{align}
We use a finite-volume formalism, so $\avg{\Uc}$ is the average of the
conserved state in a zone, $\avg{\Advs{\Uc}}^m$ is the average
advective term in a zone at time $t^m$, and $\avg{\Rbs{\Uc}}^{k+1}$ is
the average of the reactive source at time $t^{m+1}$.  The second
superscript on each term, $(k)$ or $(k+1)$, is the iteration.  This
update is an implicit equation for the new state,
$\avg{\Uc}^{{m+1},(k+1)}$.  The last term in the update is an integral
over the sources constructed from the previous iteration's values at
each time node.  Each iteration of the SDC method increases the
temporal order of accuracy by one, up to the order of accuracy with
which the integral is constructed.

\castro\ implements both a second-order method (using a trapezoid rule
for the integral) and fourth-order method in space and time (using a
Simpsons rule for the integral and the spatial reconstruction of
\cite{mccorquodalecolella}).  We demonstrated that \castro\ achieves
the expected convergence on a wide variety of problems in
\cite{castro:sdc}.  At the moment, the method is limited to single
levels, but work is underway to extend this methodology to AMR.  For
problems where burning is imprtant and can dominate the computational
expense, we expect these new SDC method will become the preferred
integration technique in the future.
